---
title: "Prepare sample plan input"
output: rmarkdown::html_vignette
vignette: >
  #%\VignetteIndexEntry{PEMsamplr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(PEMprepr)
library(PEMsamplr)
library(sf)

```

## Develop a Sample Plan

Sample plan requires running the PEMprepr functions to acquire and prepare all base data. 
Help in doing this can be found in the [PEMprepr package](https://github.com/bcgov/PEMprepr) with a detailed vinette [here](https://github.com/bcgov/PEMprepr/blob/master/vignettes/SetUpNewAOI.Rmd)

We will continue to follow the vignette using the example Canyon Creek study area. 


1) Generate AOI for landscape variables 

```{r}
# set up the folder structure and define folder names in object fid
fid <- setup_folders("CanyonCreek")

# read in AOI
aoi_raw <- system.file("extdata", "aoi.gpkg", package ="PEMprepr")
aoi_raw <- sf::st_read(aoi_raw)

# 2) snap to adjoining watersheds to provide wider landscape context 
aoi_landscape <- PEMprepr::aoi_snap_landscape(aoi_raw)

# generate a 25m template for 
t25 <- create_template(aoi_landscape, 25)

# read in DEM data from CDED
trim_raw <- bcmaps::cded_raster(aoi_landscape)

# convert the trim to matching raster 
trim <- terra::rast(trim_raw)
trim <- terra::project(trim, t25)

# generate the covariates at landscape

outpath <- 'D:/PEM_DATA/PEMsamplr/CanyonCreek/20_sample_plan/10_standard_sample/10_input_raster/landscape_covariates'

SAGApath_dir ="C:/SAGA/saga-7.7.0_x64/"

template = terra::rast("D:/PEM_DATA/PEMsamplr/CanyonCreek/10_clean_inputs/20_covariates/25m/bec.tif")

template

create_samplr_covariates(trim, SAGApath =  SAGApath_dir,
                              output =  outpath,
                              sieve_size = 10,
                              rtemplate = template)


## generate the BEC template and then match the covariates to size
# 
# bec_sf <- sf::st_read(file.path(fid$shape_dir_1010[1], "bec.gpkg")) %>%
#   sf::st_cast(., "MULTIPOLYGON") 
# 
# bec_code <- bec_sf %>% st_drop_geometry()  %>% dplyr::select(MAP_LABEL) %>%
#     unique() 
# 
# bec_code <- bec_code %>% 
#     dplyr::mutate(bgc_unique_code = seq(1, length(bec_code$MAP_LABEL),1))
# 
# bec_sf <- dplyr::left_join(bec_sf, bec_code)
# 
# bec_vec <- terra::vect(bec_sf)
# 
# # generate a 25m raster

# bec_ras25 <- terra::rasterize(bec_vec, r25, field = "MAP_LABEL")
# 
# terra::writeRaster(bec_ras25, file.path(fid$cov_dir_1020[2], "25m", "bec.tif"), overwrite = TRUE)


```

Assess how well the covariates describe the landscape and adjust as needed. Write out the landscape variability validation layer.

```{r}

outpath <- 'D:/PEM_DATA/PEMsamplr/CanyonCreek/20_sample_plan/10_standard_sample/10_input_raster/landscape_covariates'

fileoi <- c("dah_LS.tif", "mrvbf_LS.tif", "landform_LS.tif")

filesoi <- list.files(outpath, full.names = TRUE)[list.files(outpath) %in% fileoi]

landscapes <- create_binned_landscape(filesoi)

terra::plot(landscapes)

terra::writeRaster(landscapes, file.path(outpath, "landscape_variable_validation.tif"), overwrite = TRUE)

```

Assess how well the environmental space is spread for each BGCs 

```{r}

becpath <- 'D:/PEM_DATA/PEMsamplr/CanyonCreek/10_clean_inputs/10_vector'
#becpath <-fid$shape_dir_1010[1]

bec <- sf::st_read(file.path(becpath, "bec.gpkg")) %>%
   sf::st_cast(., "MULTIPOLYGON") 

routdf <- check_bgc_landscapes(bec, landscapes )

library(ggplot2)

p1 <- ggplot2::ggplot(routdf, aes(landscape)) +
  ggplot2::geom_histogram() +
  ggplot2::facet_wrap(~MAP_LABEL)

p1

p2 <- ggplot2::ggplot(routdf, aes(landscape, fill = MAP_LABEL)) +
  ggplot2::geom_histogram() 

p2

```

