---
title: "Generate sample plan"
output: rmarkdown::html_vignette
vignette: >
  #%\VignetteIndexEntry{PEMsamplr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(PEMprepr)
library(PEMsamplr)
library(sf)

```

## Generate sample plan : part 2

Sample plan requires running the PEMprepr functions to acquire and prepare all base data. 
Help in doing this can be found in the [PEMprepr package](https://github.com/bcgov/PEMprepr) with a detailed vinette [here](https://github.com/bcgov/PEMprepr/blob/master/vignettes/SetUpNewAOI.Rmd)

We will continue to follow the vignette using the example Canyon Creek study area. 


```{r}
#  define folder names in object fid
fid <- setup_folders("CanyonCreek")


## prepare base layers for cost layer

#vec_dir = fid$shape_dir_1010[2]
vec_dir = "D:\\PEM_DATA\\PEMsamplr\\CanyonCreek\\10_clean_inputs\\10_vector" 

dem = terra::rast(file.path(fid$cov_dir_1020[2], "25m", "dem_preproc.sdat"))

costprep <- prep_cost_layers(vec_dir, dem)



