---
title: "Generate sample plan"
output: rmarkdown::html_vignette
vignette: >
  #%\VignetteIndexEntry{PEMsamplr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(PEMprepr)
library(PEMsamplr)
library(sf)

```

## Generate sample plan : part 2

Sample plan requires running the PEMprepr functions to acquire and prepare all base data. 
Help in doing this can be found in the [PEMprepr package](https://github.com/bcgov/PEMprepr) with a detailed vinette [here](https://github.com/bcgov/PEMprepr/blob/master/vignettes/SetUpNewAOI.Rmd)

We will continue to follow the vignette using the example Canyon Creek study area. 


```{r}
#  define folder names in object fid
fid <- setup_folders("CanyonCreek")


## prepare base layers for cost layer

#vec_dir = fid$shape_dir_1010[2]
vec_dir = "D:\\PEM_DATA\\PEMsamplr\\CanyonCreek\\10_clean_inputs\\10_vector" 

dem = terra::rast(file.path(fid$cov_dir_1020[2], "25m", "dem_preproc.sdat"))
#sort(unique(round(terra::values(dem),0)))


costprep <- prep_cost_layers(vec_dir, dem)

terra::plot(costprep)

# Define start points 
# 1) used pre defined the start location using nearest town

cities <- st_read(file.path(vec_dir, "major_towns_bc.gpkg"))
nearest_town = "Granisle"
start <- cities[cities$NAME == nearest_town,"NAME"]%>% 
  as("Spatial")

cost <- create_cost_layer(costprep, start)

terra::plot(cost)

# note if you get an error here - be sure the town name is located on the road network of might fail. 

# Check how the cost is distributed per 








# generate cost penalty 

#vec_dir = fid$shape_dir_1010[2]
vec_dir = "D:\\PEM_DATA\\PEMsamplr\\CanyonCreek\\10_clean_inputs\\10_vector" 
cost <- cost 
dem = terra::rast(file.path(fid$cov_dir_1020[2], "25m", "dem_preproc.sdat"))
costval = 3000

final_cost <- create_cost_penalty(vec_dir = "D:\\PEM_DATA\\PEMsamplr\\CanyonCreek\\10_clean_inputs\\10_vector", 
                                  cost = cost, 
                                  dem = dem , 
                                  costval = 3000)
terra::plot(final_cost)




