---
title: "Generate heli sample plan"
output: rmarkdown::html_vignette
vignette: >
  #%\VignetteIndexEntry{PEMsamplr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}

library(PEMprepr)
library(PEMsamplr)
library(sf)

```

## heli sample Planing 

Helicopter sample planning requires a differnce cost layer to be generated in order to accurately assess costs. 

Drop locations should be defined prior to starting this process. These can eb determined using google earth or equivalent imagery to identify potential landing sites. 

# Create a heli cost layer 


```{r}
# set output folder for clhs plans
clhs_outpath <- fid$samplingplan_clhs[2]
ifelse(!dir.exists(clhs_outpath), dir.create(clhs_outpath, recursive = TRUE), FALSE)

#vec_dir = fid$shape_dir_1010[2]
vec_dir = "D:\\PEM_DATA\\PEMsamplr\\CanyonCreek\\10_clean_inputs\\10_vector" 

dem = terra::rast(file.path(fid$cov_dir_1020[2], "25m", "dem_preproc.sdat"))
#sort(unique(round(terra::values(dem),0)))

# prepare cost layer
costprep <- prep_cost_layers(vec_dir, dem, heli = TRUE)
terra::plot(costprep)


# read in the heli drop points
helidrop <- st_read(file.path(vec_dir, "helidrop.gpkg"))%>% 
  as("Spatial")

# create cost layer 
cost <- create_cost_layer(costprep, helidrop)
cost <- terra::rast(cost)
names(cost) <- "cost"
terra::plot(cost)


# read in the landscape covariates 
landscape_dir <- 'D:/PEM_DATA/PEMsamplr/CanyonCreek/20_sample_plan/10_standard_sample/10_input_raster/landscape_covariates'

fileoi <- c("dah_LS.tif", "mrvbf_LS.tif", "landform_LS.tif")

filesoi <- list.files(outpath, full.names = TRUE)[list.files(outpath) %in% fileoi]

all_cov <- terra::rast(filesoi)
all_cov <- c(all_cov, cost)


# read in bec labels 
bec_dir = "D:\\PEM_DATA\\PEMsamplr\\CanyonCreek\\20_sample_plan\\10_standard_sample\\10_input_raster\\exclusion" 

boi <- list.files(bec_dir, pattern = ".tif")
# note for heli sampling in most cases it will only be a specific Bec unit and will not need to use all options

b<- boi[2]

boi_mask <- terra::rast(file.path(bec_dir, b)) 
names(boi_mask) = "mask"
bname <- gsub("_exclude_mask.tif", "", b)
  
sample_layers_masked <- c(all_cov, boi_mask) %>%  
    terra::mask(boi_mask) 
sample_layers_masked <- sample_layers_masked[[1:4]]
  
# create 10 different sample plans
  
for(rot in 1:5){ 
  
    sample_points <- create_clhs(all_cov = sample_layers_masked, 
                                 num_slices = 2, 
                                 to_include = NULL, 
                                 n_points = 5 , 
                                 min_dist = 1000,
                                 num_sample = 5000000)
    
    sf::st_write(sample_points, file.path(clhs_outpath, paste0(bname,"_clhs_heli_sample_",rot,".gpkg")),delete_dsn = TRUE)
    
  }
  


```












# select the clhs with the lowest sample cost

1) Using simple cost metrics 
2) Using vehicle routing problem 

```{r}
library(foreach)
# read in all the points per bgc, collate and assign value based on travel cost
 
ftemp <- list.files(clhs_outpath, pattern = ".gpkg$", full.names = TRUE)

all_samples <- foreach(fs = 1:length(ftemp), .combine = "rbind" ) %do% {
  
  ff <- ftemp[fs]
  layers <- st_layers(ff)$name
  
  sample_points_all <- foreach(l = 1:length(layers), .combine = rbind) %do% {
    ll = layers[l]
    sptemp <- st_read(ff, layer = ll)
    sptemp <- sptemp %>%
      dplyr::mutate(clhs_repeat = basename(ff)) %>% 
      dplyr::mutate(bgc = stringr::str_extract(layers, "[^_]+"))
 
    sptemp
  }
  
}

# summarise the costs per sample plan
repsum <- all_samples %>%
  sf::st_drop_geometry() %>%
  dplyr::select(clhs_repeat, bgc, cost)

repsum <- repsum %>%
  dplyr::group_by(clhs_repeat, bgc) %>%
  dplyr::mutate(tcost = sum(cost)) %>%
  dplyr::select(-cost) %>%
  dplyr::distinct()

# plot the total costs by subzone
p1 <- ggplot(repsum, aes(y = tcost, x =  clhs_repeat )) +
  geom_point()+
  facet_wrap(~bgc, scales = "free_x")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
p1

```

2) Alternatively run Vehicle routing problem 

```{r}
# set up start location
vec_dir = "D:\\PEM_DATA\\PEMsamplr\\CanyonCreek\\10_clean_inputs\\10_vector"
cities <- st_read(file.path(vec_dir, "major_towns_bc.gpkg"))
nearest_town = "Granisle"
start <- cities[cities$NAME == nearest_town,"NAME"]

## THis needs work to complete vrp 

```



3) Generate the additional transect information 




