---
title: "05_Import_field_data"
author: "G. Perkins"
date: "2022-12-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import and Clean Field data 

Once the field data has been collected, we can preprocesses. 
Depending on the size of the study area or areas, we reccomend storing the track and point data in a spatial file or database. 

In this example we will process a group of spatial files - either shapefiles or geopackages and tidy them to be put into a central database/spatial database.

In this example we will use some test data

```{r libraries}

load_all()
library(PEMprepr)
library(sf)

```


```{r pressure, echo=FALSE}
# set up the folder names
fid <- setup_folders("CanyonCreek")

# location of the transect layout (sample plan)
trans_input <- fid$sampleplan_final_transect[2]
trans_output <- fid$trainpts_201040[2]
  
# location of the raw field data
rawdat <- fid$trainpts_transect[2]

# output location for cleaned data
cleandat <- fid$trainpts_maps[2]
  
```

# Create a consolidated transect layout 

We will firstly create a basic layer with all the transects names consolidated for each BEC. This will be used as a template for assigning incidental or transect data points 


```{r}

# test this on multiple locations 
# 
inpath <- "D:\\PEM_DATA\\BEC_DevExchange_Work/BoundaryTSA_AOI/2_sample_design/stage1_StudyDesign/transect_layout"
# 
# inpath <- "D:\\PEM_DATA\\BEC_DevExchange_Work\\BoundaryTSA_AOI\\2_sample_design\\stage1_StudyDesign\\transect_layout"
# 
# inpath <- "D:\\PEM_DATA\\BEC_DevExchange_Work\\WilliamsLake_AOI\\2_sample_design\\stage1_StudyDesign\\transect_layout"

transect_layout <- generate_transectlayout(inpath)

transect_layout_buf <- sf::st_buffer(transect_layout, 10)

st_write(transect_layout, file.path(trans_output, "transect_layout_s1.gpkg"), delete_layer = TRUE)

```


# import and clean field data

```{r}
# 
rawdat <- "D:\\PEM_DATA\\BEC_DevExchange_Work\\BoundaryTSA_AOI\\2_sample_design\\stage1_StudyDesign\\transect_data\\2019"
# 
# rawdat <- "D:\\PEM_DATA\\BEC_DevExchange_Work\\WilliamsLake_AOI\\2_sample_design\\stage1_StudyDesign\\transect_data"

#rawdat <- "D:\\PEM_DATA\\BEC_DevExchange_Work\\KootInvCran_AOI\\2_sample_design\\stage1_StudyDesign\\transect_data\\final_contractors"


library(foreach)

points <- format_fielddata(rawdat, transect_layout_buf)
#points <- all_points

# covert to lines 
processed_transects <- make_lines(GPSPoints = points, 
                                  Transects = transect_layout_buf, 
                                  method = "pts2lines",  
                                  tBuffer = 20, PROJ = 3005) %>%
  dplyr::select(-TID, -ID)


st_write(processed_transects,  file.path(cleandat, "proc_s1_transects_b2020bcgov.gpkg"), 
         delete_layer = TRUE)

sf::st_write(points, file.path(cleandat, "s1_points_b2020bcgov.gpkg"), delete_layer = TRUE)


# format tracklog
#rawdat <- "D:\\PEM_DATA\\BEC_DevExchange_Work\\KootInvCran_AOI\\2_sample_design\\stage1_StudyDesign\\transect_data\\final_contractors\\_dev\\"

tracks <- format_tracklog(rawdat, transect_layout_buf)

sf::st_write(tracks, file.path(cleandat, "s1_tracklog_b2020bcgov.gpkg"), delete_layer = TRUE)

```


