---
title: "06_Attribute_field_data.Rmd"
author: "G. Perkins"
date: "2023-03-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Mapping resolution and mapping units 

In many of the site series we can roll up finer classification into broader groups. To do this we need to use a build to match field calls to map units. 

In this case we have a pre-populated legend. 

Once the final level of modelling is decided we convert points to line segments, and then extract all points of the raster size in whcih we will model with, default of 5m. 



## Prepare and attribute all training data points 

1) convert lines to points 
2) extract neighbours 
3) attribute points 
4) select mapkey to use in models


```{r libraries}

load_all()
library(PEMprepr)
library(sf)
library(terra)

```


Create the set of folders is not already running from previous scripts

```{r}
# set up the folder names
fid <- setup_folders("CanyonCreek")

#map.key  <- read.csv(file.path(fid$AOI_dir[2], "DateCreek_MapUnitLegend.csv"), stringsAsFactor = FALSE)

## clean points 
#cleandat <- fid$trainpts_maps[2]
cleandat <- "D:\\PEM_DATA\\DateCreek_AOI\\DateCreek_AOI\\20_sample_plan\\10_standard_sample\\40_transect_data\\clean_field_data"

res_folder = "5m"

#location of raster base on which modelling will be applied
#covdir <- fid$cov_dir_1020[2]
covdir <- "D:\\PEM_DATA\\DateCreek_AOI\\DateCreek_AOI\\10_clean_inputs\\20_covariates"

# location of attributed and output data 
#outdat <- fid$training_data_1030[2]
outdat <- "D:\\PEM_DATA\\DateCreek_AOI\\DateCreek_AOI\\10_clean_inputs\\30_trainingdata"
```

#1) Convert lines to points 


We can choose to extract all adjoining points for each training point. 
This process is done to overcome spatial missregulation errors. 


```{r}
# read in the raster template used for modelling (i.e 5m resolution)
trast <- terra::rast(file.path(covdir, res_folder,"template.tif"))

#tpoints <- st_read(file.path(cleandat, "s2_points.gpkg"))

processed_transects <- st_read(file.path(cleandat, "proc_s1_transects.gpkg"))

if("ID" %in% colnames(processed_transects) == FALSE){
  print ("adding ID column")
  processed_transects <- processed_transects %>% 
    mutate(ID = seq(1, length(processed_transects$order), 1))
}

processed_transects_id <- st_drop_geometry(processed_transects)

# transv <- terra::vect(file.path(cleandat, "proc_s1_transects_datecreek.gpkg"))

lBuff <- processed_transects %>% 
    sf::st_buffer(., dist = 2.5, endCapStyle = "FLAT", joinStyle = "MITRE") %>%
    sf::st_cast(.,"MULTIPOLYGON")

# convert to spatVect
transv <- terra::vect(lBuff )

# extract the XY values of raster cells where crossed lines
xys <- terra::extract(trast,transv, xy = TRUE )

# add the additional points back to full dataset
raster_points_xy <- xys %>% 
  st_as_sf(coords = c("x", "y"), crs = 3005) %>%
  merge(processed_transects_id) %>%
  select(-ID, -template)

# add slice and tid (transect id)

allpts <-raster_points_xy %>%
    mutate(tid = tolower(gsub("_[[:alpha:]].*","", transect_id))) %>%
    mutate(slice = sub('.*(?=.$)', '',gsub("\\..*","", tid), perl=T))
  
# write out? 

```


Add the adjoining cells 

```{r}

tpoints_ne <- add_neighbours(allpts,trast)

```

Attribute all the data 

```{r}

head(tpoints_ne)

allrasts <- file.path(covdir, res_folder)

#cov_dir = allrasts

allpts <- attribute_points(tpoints_ne, allrasts) 

st_write(allpts, dsn = file.path(fid$training_data_1030[2], "allpts.gpkg"), delete_layer = TRUE)

```

