---
title: "Field training point summary report"
output: html_document
params:
  tpts: tpts
  trans: trans
  out_dir: out_dir
---

This report is a breakdown of training point data and line segments. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=FALSE,
                      warning = FALSE, message = FALSE,
                      results = 'show',  fig.width = 14,
  fig.height = 10,eval = TRUE)  ## flag eval = false for quick text edits
```


```{r global_options, include=FALSE}

## Load library
#library(data.table)
#library(rmarkdown)
#library(knitr)
#library(dplyr)
#library(ggplot2)

tpts <- params$tpts
trans <- params$trans
out_dir <- params$out_dir

```


## Data type
### Points

Points can be allocated into sampling stage types (s1,s2) or where points are collected outside of standard transect (incidentals). 

```{r format raw point data, echo = FALSE}
ptsdf <- tpts %>%
  sf::st_drop_geometry()

pttype <- ptsdf  %>% 
  dplyr::select(data_type) %>%
  group_by(data_type)%>%
  dplyr::mutate(count = n()) %>%
  dplyr::distinct()

knitr::kable(pttype)

```

### Length of transect segments

```{r, echo = FALSE}
all_trans <- trans %>%  
    mutate(length = st_length(.)) %>%
    st_drop_geometry() 

# total type of points with primary and secondary 
trans_type_p <- all_trans %>%
  dplyr::group_by(data_type) %>%
  dplyr::summarise(primary = sum(length, na.rm = TRUE))

trans_type_s <- all_trans %>%
  filter(!is.na(mapunit2)) %>%
  group_by(data_type) %>%
  dplyr::summarise(secondary = sum(length, na.rm = TRUE))

trans_type <- left_join(trans_type_p, trans_type_s) %>%
  mutate(pc.sec = round((secondary/primary * 100),1))

trans_type_long <- left_join(trans_type_p, trans_type_s) %>%
  mutate(pc.sec = secondary/primary * 100)

trans_types <- unique(trans_type$data_type)

```

A total length of `r round(sum(trans_type$primary),1)` meters was surveyed during transect collection. This include `r round(length(trans_type$data_type),1)` data types : `r trans_types`.

Alternate calls were made for `r round(sum(trans_type$secondary, na.rm = TRUE),2)` meters, with an average of `r round(sum(trans_type$secondary, na.rm = TRUE)/sum(trans_type$primary, na.rm = TRUE)*100, 1) ` % of all calls.

```{r, echo = FALSE}
knitr::kable(trans_type)
```

## Transect summary 
The number of sites and slices per BGC. A full list of transect points per trasnect are show in Appendices below

```{r, echo = FALSE, fig.width = 8, fig.height=5}
# no of slices
nopts <- ptsdf %>% 
  dplyr::select(transect_id) %>%
  group_by(transect_id) %>%
  dplyr::mutate(count = n()) %>%
  dplyr::distinct() %>% 
  arrange(transect_id)

trans_slice_site <- nopts %>%
  dplyr::rowwise() %>%
  mutate(bgc = unlist(strsplit(transect_id, "_"))[1]) %>%
  mutate(slice_site = unlist(strsplit(transect_id, "_"))[2]) %>%
  mutate(slice = stringr::str_sub(slice_site, 1,1)) %>%
  mutate(site = stringr::str_sub(slice_site, 3,3))       

trDat_slices <- trans_slice_site %>%
  dplyr::group_by (bgc) %>%
  dplyr::summarise(n.transect = length(unique(transect_id)),
                   n.sites = length(unique(slice_site))) %>% 
  tidyr::pivot_longer(cols = c("n.transect", "n.sites"), names_to = "type", values_to = "number")

ggplot2::ggplot(trDat_slices, aes(bgc, number, fill = type)) +
  geom_bar(stat = "identity", position = "dodge") + 
  theme(axis.text.x = element_text(angle = 90))+
  theme_minimal()+
  colorspace::scale_fill_discrete_sequential(palette = "Light Grays")


```

### Field Observers

```{r observer summary, echo = FALSE}
# noobs <- ptsdf  %>% 
#   dplyr::select(observer) %>%
#   group_by(observer) %>%
#   mutate(count = n()) %>%
#   distinct()
# 
# knitr::kable(noobs)

# total type of points with primary and secondary 
ptob_type_p <- ptsdf %>%
  filter(!is.na(mapunit1)) %>%
  st_drop_geometry() %>%
  dplyr::group_by(observer) %>%
  dplyr::summarise(primary = n())

ptob_type_s <-  ptsdf%>%
  st_drop_geometry() %>%
  filter(!is.na(mapunit2)) %>%
  group_by(observer) %>%
  dplyr::summarise(secondary = n())

trans_type <- left_join(ptob_type_p, ptob_type_s,by = join_by(observer)) %>%
  mutate(pc.sec = round((secondary/primary * 100),0))


knitr::kable(trans_type)



```

There were `r length(noobs$observer)` observers recording data, including unnamed recorders (NA). The table below shows the total calls and the proportion of calls which also had a secondary call. 

```{r, include = FALSE}
# total type of points with primary and secondary 

pt_type_p <- ptsdf  %>%
  group_by(data_type) %>%
  dplyr::summarise(all.calls = n())

pt_type_s <- ptsdf %>%
  filter(!is.na(mapunit2)) %>%
  group_by(data_type) %>%
  dplyr::summarise(sec.calls = n())

pt_type <- left_join(pt_type_p, pt_type_s) %>%
  mutate(percent.sec = sec.calls/all.calls * 100) 

#pt_types <- unique(pt_type$data_type)

knitr::kable(pt_type)

```
## Mapunits per BGC

```{r, echo = FALSE}
ppts <- ptsdf %>%
  mutate(mapunit2 = ifelse(mapunit2 == "", NA, mapunit2),
         mapunit1 = ifelse(mapunit1 == "", NA, mapunit1)) %>%
  group_by(mapunit1, mapunit2) %>%
  dplyr::mutate(count = n()) %>%
  ungroup()%>%
  filter(is.na(mapunit2))

ppts <- ppts %>%
  dplyr::arrange(mapunit1) %>%
  dplyr::rowwise() %>%
  mutate(bgc = unlist(strsplit(mapunit1, "/"))[1]) %>%
  mutate(bgc = ifelse(stringr::str_detect(mapunit1, "/"), bgc, "non-forest")) %>%
  ungroup()

ppts <- ppts %>%
  dplyr::select(mapunit1, bgc) %>% 
  group_by(mapunit1)%>%
  dplyr::mutate(count = n()) %>%
  dplyr::distinct()

pptsbec <- ppts %>%
  dplyr::group_by(bgc) %>%
  dplyr::summarise(bgc_freq = sum(count)) 
 # mutate(prop = round(count/sum(count),3))

ppts <- ppts %>%
  dplyr::left_join(pptsbec) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(prop = round(count/bgc_freq,3)*100) %>%
  arrange(bgc)

#knitr::kable(ppts)

bgc_list <- unique(ppts$bgc)

```

```{r, echo = FALSE, fig.width = 8, fig.height=5}
# plot the freqency 

for(b in bgc_list) {
 # b = unique(pt_total$bgc)[3]
  pt_total_bgc <- ppts %>%
    filter(bgc == b)
  
  p <- ggplot(pt_total_bgc , aes(x = reorder(mapunit1, prop), y = prop)) +
  geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90))+
  #theme_pem()+
  colorspace::scale_fill_discrete_sequential(palette = "Light Grays")

  labs(title = b, y = "mapunit")
  
  print(p)
  
}
```

## Appendices

The number of points per transects. Note this is the raw points and not the processed transects so point numbers will vary. 

```{r}
knitr::kable(nopts)
```

